Using Bayesian Networks to Predict Change Sets in an IT System

### 2010-12-10 cs886-sarah/Paper.tex
semi final draft isA

\documentclass{article}

% For figures
\usepackage{graphicx}

% For citations
\usepackage{natbib}

% For algorithms
\usepackage{algorithm}
\usepackage{algorithmic}
\usepackage{url}

% As of 2010, we use the hyperref package to produce hyperlinks in the
% resulting PDF.  If this breaks your system, please commend out the
% following usepackage line and replace \usepackage{icml2010} with
% \usepackage[nohyperref]{icml2010} above.
\usepackage{hyperref}

% Packages hyperref and algorithmic misbehave sometimes.  We can fix
% this with the following command.
\newcommand{\theHalgorithm}{\arabic{algorithm}}
\usepackage[accepted]{icml2010}

% The \icmltitle you define below is probably too long as a header.
% Therefore, a short form for the running title is supplied here:
\icmltitlerunning{Title}

@@

\twocolumn[
\icmltitle{Using Bayesian networks to predict changes}

\icmlauthor{Sarah Nadi}{snadi@uwaterloo.ca}
\icmladdress{University of Waterloo,
            Waterloo, ON, Canada}

\vskip 0.3in
]

\begin{abstract} 
STILLLL
\end{abstract}

@@

Change is inevitable in any Information Technology(IT) system. New features are added, different configurations are used, upgrades are introduced and new
software and hardware are added. In a large system, such changes can cause other parts of the system to malfunction without the ability of the analyst to
foresee this side effect. Accordingly, we need to ensure that we consider the impact of all changes before they are implemented, and make sure we adjust the
potentially affected components accordingly while implementing the change. This leads to the process of change set detection which is finding all the
components of the system that need to be included in your change set where the change set is the set of components that need to be changed to avoid any side
effects.

Many of the large organizations with large and sophisticated IT systems employ a Configuration Management Database (CMDB) to help them manage these systems.
The CMDB keeps track of all the components in a system, how they are related, as well as all the changes that have occurred to them. Any component in the CMDB
is called a Configuration Item (CI). In previous work~\cite{nadi2010}, we mined the CMDB repository for historical co-changes, and used the mined correlations
to predict change sets. That is, given an initial CI that is going to change, we predict what other CIs might need to be changed as well based on its change
history with other CIs. We obtained really good results in terms of recall and precision (69.8\% and 88.5\% respectively).

In this paper, we wish to examine the same problem, but from a Bayesian perspective. Instead of looking at pairwise historical co-changes of CI, we would like
to consider all the changed CIs in the past to deduce the causality relations between them. That is, what is the probability that if A changes, B changes as
well. In that case, A would be a parent to B in a Bayesian network. Accordingly, we explore the different ways a Bayesian network can be constructed from the
data we have, and then test the predictions produced by querying this network to examine the obtained recall and precision. We hope that we can obtain better
results using Bayesian networks since they are not limited to pairwise comparisons.

The rest of this paper is organized as follows. BLA BLA BLA

\section{Background}

\subsection{CMDBs and Change Sets}
A Configuration Management Database (CMDB)~\cite{itil} is useful in Enterprise IT Management (EITM) since it provides information about the various critical
components in a
system including hardware, software, and services provided by the company. It records the configuration of these items, their change history, their incident
history, as well as the relationships between them. Each item stored in the CMDB is referred to as a Configuration Item (CI). Figure~\ref{fig:cmdbExample} shows
an example CMDB to illustrate the concepts of CIs and relationships.

A CMDB provides a basis for decision making processes such as Incident Management, Change Management, etc.  In this paper, we focus on the process of Change
Management, and in particular, on the problem of change set detection. A \textit{change} is the addition, modification, or removal of anything that could affect
on IT services. A poorly planned change may lead to a fault in the system. Accordingly, when one wants to change one CI is the system, other CIs that might
need to be changed as well must be correctly identified. The set of CIs that will need to be modified for the change to be complete is called a change set.

@@

@@

\subsubsection{Weka}

WEKA~\cite{hall2009weka} is a toolbox for machine learning. It provides different algorithms for classification and clustering as well as other machine
learning problems. It also provides Bayesian learning techniques for classification problems. Additionally, it provides explorations tools for Bayesian
Networks, and allows the user to learn the structure and CPTs of a Bayesian Network. We will mainly be using WEKA for learning the CPTs in a fixed
structure network using the SimpleEstimator algorithm~\cite{written2005data} implemented there. This simple estimates the probability values on edges based on
the frequency values of each variable given its parents in the training set.

\subsubsection{Banjo}
Banjo~\cite{banjotool} is a tool written in Java which infers the structure of a Bayesian network given training data. It has two different search algorithms:
Greedy and Simulated Annealing. For either of these search algorithms, it has two methods of proposing a new edge. The first is the ``proposeRandomLocalMove''
which basically proposes a random addition or removal of an edge. The other is ``proposeAllLocalMoves'' which proposes all possible moves, and only keeps the
best one (?). Banjo uses the BDe metric to compute a network's score.

@@

Mirarab et al.~\cite{mirarab2007} investigate the same problem as our work. However, their work is on the level of source code changes. They build three
different Bayesian Networks, one that is based on package and class dependency information (static relationships), one which is dependent on historical
co-changes, and one which uses both. For the first graph, the initial structure is essentially ``given'' according to the static dependencies, and then the CPTs
are learnt using the importance sampling algorithm proposed by Changhe and Marek~\cite{yuan2003importance}. The way static dependencies are defined in their
case is specific to Java. The third one is essentially the first graph, but updated using the historic change information according to the Expectation
Maximization (EM) algorithm~\cite{dempster1977maximum}. The second was solely based on historic information where the network is build using a greedy structure
 learning algorithm~\cite{friedman1996learning}. They did some preprocessing to their data such as filtering out large changes (with more than 30 elements
changed at once) since this was probably an insignificant change.

Zhou et al.~\cite{zhou2008} try to answer a slightly different problem. They do not only look at the probability of other elements changing given a specific
element, they also add features such as authors, change significance levels etc. and try to predict if two elements are co-changes or not accordingly. Thus,
their problem is more of a classification problem where given two elements, and some observed features they try to determine the class as co-changes or not.
They use the K2 algorithm proposed by Cooper et. al~\cite{cooper1992bayesian} to estimate the structure of the Bayesian network, and use the SimpleEstimator
algorithm built in WEKA~\cite{witten2005data}.

\section{Constructing the Bayesian Network}
\label{sec:modelsused}

Before constructing the Bayesian Network, we first had to process and prepare the data we have. Then, in order to construct a Bayesian network to use for
predictions there are two steps involved. First, determining the structure of the actual network, and then estimating the Conditional Probability Tables
(CPT). Section~\ref{sec:data} first explains the data set available to us, and the data preprocessing involved before building the network.
Section~\ref{sec:structure} then explains the different techniques we experimented with to build the network structure. Section~\ref{sec:cpt} shows the last
step to build the network which is estimating the CPTs. Figure~\ref{fig:process} shows this overall process.

@@

\subsubsection*{Data Size}
The original data set is from three years, and has 7,999 distinct CIs, and 27,305 change orders. This amount of data was infeasible to work with as no tool
could handle such a large amount of variables in a Bayesian network. For the purposes of this project, which is mainly to experiment with Bayesian techniques
for change set prediction,
it is sufficient to choose a small representative subset of the data. Accordingly, in order to be able to test things properly, we used observations from three
months data from January 1, 2008 to March 31, 2008 to build the model. However, even in such a short period, there were already 2,841 distinct CIs appearing and
2,229 observations (i.e. change orders). Therefore, we needed to perform further data preprocessing.

\begin{figure}[!t]
\centering
\includegraphics[width=7cm]{graphics/constructingmodel.png}
\caption{Building the Bayesian Network}
\label{fig:process}
\end{figure}

First, we removed all CIs that have changed less than 12 times within this time frame  (i.e were associated with 12 different change orders in our training
observations). There is no particular reason for choosing 12 as a cut off. It simply gave a feasible data set to deal with. This yielded 120 distinct CIs. Then,
in order to slightly increase our variable space to include other related CIs, we found all the parent CIs related to these 120 CIs from the CMDB perspective.
However, we ignored three common, and not extremely meaningful relations, which are  ``supports'', ``is location for'', and ``backs up''. Adding the related
parent CIs, we now had a set of 241 CIs. However, some of the added parent CIs may not be in the original set of 2,841 CIs. Accordingly, we just kept CIs tht
appeared more than 12 times or were in the set of related CIs. This provided us our final data set of 170 CIs which we use throughout our models for fair
comparison. Additionally, filtering out CIs meant filtering out some of the observations that did not have any CIs satisfying our criteria. This led to us
having 1,305 observations instead of 2,229 which was a more manegeable set. Accordingly, our training data set consisted of 170 variables (CIs), and 1,305
observations (change sets).

FIGURE FOR DATA PROCESSING

@@

In the CMDB, a Change Order has several fields including the requester, the assignee, the start date of the change, the description and the change set field
(called the `Configuration items' fields). Of the fields in a change order, we use only the change set field. The advantage of only using this one field is that
change sets are easy to extract and easy to understand. In terms of an observation, CIs were either marked as ``true'' or ``false'' indicating whether they
have changed or not. Each observation initially consisted of all the in our data set marked as ``false''. For each change order in the training set, we would
mark each CI appearing in the ``Configuration Items'' field as true, while all those not appearing were left as ``false''. Therefore, the set of training
data consisted of an observation entry for each change order with the CIs marked as ``true'' or ``false'' accordingly.

\subsection{Network Structure}
\label{sec:structure}
There were different ways in which we could estimate the structure of the network, and so we built a Bayesian Network using each technique.

\subsubsection{Model 1 (Banjo)}
\label{sec:model1}

For the first model, we used Banjo to learn the structure of the Bayesian network from the training data. We use the Greedy searching algorithm implemented
there. Banjo produced five different top-scoring networks (having the same score), and we simply chose one of them as the network to be used.

\subsubsection{Model 2 (CMDB Relationships)}

For the second model, we used the relations existing in the CMDB to infer the structure of the network (that is place the edges between the nodes). We used two
variations for this. The first was to follow the direction of the relationship edges in the CMDB. That is if there is an edge from A to B, we will place an
edge in the Bayesian network from A to B (we will call this model 2A). The second was to reverse the direction of the relationship edges in the CMDB. That is,
if there is an edge from A to B, we will place an edge in the Bayesian network from B to A (we will call this model 2B). In both cases, however, we chose to
ignore some relationship edges (WHY, TRY WITHOUT). There was one problem, however, with the generated network. Two of the CIs had more than 15 parents to them
which means that their CPTs will be intractable to compute. For those two CIs, we simply removed all their parents to make the computation tractable.

\subsubsection{Model 3 (Missing Data)}

This model is the same as the second model, except that we experimented with the missing data feature in Weka. Weka allows the user to specify missing values
for any variable by simply putting a '?' instead of its value. Therefore, instead of putting the CIs that did not appear in a change order as false, we put a
'?' in their place. This seemed closer to practice, because in reality, we are not sure whether this CI actually changed or not. Therefore, this model also has
two parts: model 3A and model 3B where the first uses the CMDB relationships in their same direction, and the second uses the reverse direction.

@@

However, for all
models, we used the SimpleEstimator algorithm~\cite{written2005data} built in Weka to calculate the
CPTs. After setting the data set to be the observations in our training set, Weka
learned the CPTs for the network produced by Banjo. This complete network was then saved as an BIF XML file to be used by JavaBayes for general inference.

\section{Experiment Setup}

\begin{algorithm}[tb]
   \caption{Generating Predictions using the Bayesian Network}
   \label{alg:simulation}
\begin{algorithmic}
   \STATE {\bfseries Input:} changeOrder
    \STATE {\bfseries Input:} BayesianNetwork
    \STATE {\bfseries Input:} threshold
    \STATE Set initialCI = first CI in changeOrder
    \STATE Set occurredSet = changeOrder - initialCI
    \STATE Initialize observedSet = {firstCI}
    \STATE Initialize predictedSet = {}
   \REPEAT  
    \STATE Initialize newPredictions = {}
    \FOR{$node$ {\bfseries to} $BayesianNetwork$}
    \STATE posterior = perform inference using Banjo
    \IF{$posterior > threshold$} 
      \STATE Add node to newPredictions   
    \ENDIF
    \ENDFOR

  \FOR{$prediction$ {\bfseries to} $newPredictions$}
    \IF{$prediction  occurredSet$}
      \STATE Add prediction to observedSet
    \ENDIF
    \ENDFOR
\STATE Add newPredictions to predictedSet

   \UNTIL{$newPredictions$ is $empty$}
\end{algorithmic}
\end{algorithm}

@@

The change set detection process provided is an iterative, collaborative process between the tool and the analyst. When the tool suggests CIs, the analyst can
accept or reject these CIs (i.e add them to the change set or not). Based on the CIs added to the change set, the analyst can ask the tool for more
suggestions. This is along the lines of ``Now, that I am also going to change these CIs, what else do I need to change?''.

To simulate this process for our experiments, we do the following. At this point, we have the Bayesian network ready, and we would like to perform Inference.
More formally, given that a CI will change (our observation), we want to infer the probability that the other CIs in the network might change as well.
Unfortunately, neither of the two tools previously used provide an Bayesian inference engine. We, therefore, use JavaBayes in this step since it accepts the
same ARFF format used by Weka. We using a one month testing set where we try to predict all the change orders in April 2008. There was a total of
883 change orders in that month.

For each change order, we would take the first CI as the initial CI to change, then we would set that as an observed node, and update
the beliefs of all the nodes (CIs) in the network using JavaBayes. We would then loop on all the updated CIs, and add those that match our threshold criteria to
the predicted change set. To simulate a real life scenario, we then checked which of these predicted CIs actually lies in the target change set we are trying to
predict. This is similar to an analyst accepting CIs into their change set. These common CIs would then also be marked as observations so that we can predict
now that I'm going to change these accepted CIs as well, what else do I need to change. Again, we would calculate the posterior probability, and continue doing
so until there are no more common CIs. All the CIs that match the threshold criteria (whether accepted by the analyst or not) are part of the predicted
set.

@@

We need a way to evaluate the predicted CIs. The recall and precision measures from the information retrieval field are appropriate for this type of evaluation.
Recall measures the proportion of correct CIs retrieved by the system, while precision measures the proportion of suggested CIs that are correct~\cite{van79}.

Similar to Hassan et. al~\cite{hassan2004predicting}, we define the \textit{Predicted Set} (P) as the set of all CIs DRACA suggests through the full the
iteration process (see Figure~\ref{fig:process}). We define the \textit{Occurred Set} (O) as the CIs remaining in the change set after excluding the Initial CI
provided by the analyst (i.e Change Set - Initial CI). The intersection of the predicted set and the occurred set, called $PO$, is the common CIs in both sets.
For each constructed change set, we then calculate the recall and precision values for the predictions according to the following
definitions~\cite{hassan2004predicting}:

@@

Algorithm~\ref{alg:simulation} shows this process, and how the recall and precision are calculated accordingly.

>>>>

\documentclass[10pt,twocolumn,letterpaper]{article}

\usepackage{cvpr}
\usepackage{times}
\usepackage{graphicx}
\usepackage{algorithm}
\usepackage{algorithmic}
\usepackage{url}
\usepackage{hyperref}

\cvprfinalcopy % *** Uncomment this line for the final submission

\def\cvprPaperID{****} % *** Enter the CVPR Paper ID here
\def\httilde{\mbox{\tt\raisebox{-.5ex}{\symbol{126}}}}

% Pages are numbered in submission mode, and unnumbered in camera-ready
\ifcvprfinal\pagestyle{empty}\fi

@@

\title{Using Bayesian Networks to Predict Change Sets in an IT System}

\author{Sarah Nadi\\snadi@uwaterloo.ca\\CS886 - Fall 2010\\University of Waterloo\\ Ontario, Canada}        
\thispagestyle{empty}
\maketitle

\begin{abstract} 
When changing one component in an IT system, an analyst must be careful not to negatively impact any of the other system components. Detecting the complete
change set (all those components that need to be changed to avoid any side effects) is therefore very important. Given historic information about how different
components of a system change together, we build a Bayesian network that can be used to predict this change set. Given an initial component to be changed in
the system, the posterior probability of the other components changing can be calculated using the network. In this paper, we experiment with different
techniques for building the Bayesian network based on the data available, and report on their predictive power in terms of recall and precision.
\end{abstract}

@@

Change is inevitable in any Information Technology(IT) system. New features are added, different configurations are used, upgrades are introduced and new
software and hardware are added. In a large system, such changes can cause other parts of the system to malfunction without the ability of the analyst to
foresee this side effect. Accordingly, we need to ensure that we consider the impact of changes before they are implemented, and make sure we adjust the
potentially affected components accordingly while implementing the change. This leads to the process of change set detection which is finding all the
components of the system that need to be included in your change set. The change set is the set of components that need to be changed to avoid any side
effects.

Many of the organizations with large and sophisticated IT systems employ a Configuration Management Database (CMDB) to help them manage these systems.
The CMDB keeps track of all the components in a system, how they are related, as well as all the changes that have occurred to them. Any component in the CMDB
is called a Configuration Item (CI). In previous work~\cite{nadi2010}, we mined the CMDB repository for historical co-changes, and used the mined correlations
to predict change sets. That is, given an initial CI that is going to change, we predict what other CIs might need to be changed as well. We obtained really
promising results in terms of recall and precision (69.8\% and 88.5\% respectively).

In this paper, we wish to examine the same problem, but from a Bayesian perspective. Instead of looking at pairwise historical co-changes of CI, we would like
to consider all the changed CIs, and learn the relationships between them. That is, given a set of observations about how different CIs change
together, can we deduce the relationships between them that would allow us to predict future change sets. Accordingly, we explore the different ways a Bayesian
network can be constructed from the data we have, and then test the predictions produced by querying this network to examine the obtained recall and precision.
We hope that we can obtain better results using Bayesian networks since they are not limited to pairwise comparisons.

The rest of this paper is organized as follows. Section~\ref{sec:bg} provides the background information necessary for this work. It explains the nature of
CMDBs, Bayesian networks, and the tools we used for analyzing Bayesian networks. Section~\ref{sec:constructing} then presents the techniques we used to build
our Bayesian network. This includes how we process the data, and build the different Bayesian networks accordingly. Section~\ref{sec:setup} details the
setup we use for our experiments, and Section~\ref{sec:results} provides the results obtained. Section~\ref{sec:future} provides future work that can be done
to improve the obtained results. Section~\ref{sec:concl} concludes this paper.

\section{Background}
\label{sec:bg}

\subsection{CMDBs and Change Sets}
The repository of information we are using is the Configuration Management Database (CMDB)~\cite{itil}. The CMDB is useful in Enterprise IT Management (EITM)
since it provides information about the various critical components in a system including hardware, software, and services provided by the company. It records
the configuration of these items, their change history, their incident history, as well as the relationships between them. Each item stored in the CMDB is
referred to as a Configuration Item (CI). Figure~\ref{fig:cmdbExample} shows an example CMDB to illustrate the concepts of CIs and relationships.

A CMDB provides a basis for decision making processes such as Incident Management, Change Management, etc.  In this paper, we focus on the process of Change
Management, and in particular, on the problem of change set detection. A \textit{change} is the addition, modification, or removal of anything that could affect
on IT services. A poorly planned change may lead to a fault in the system. Accordingly, when one wants to change a CI is the system, other CIs that might
need to be changed as well must be correctly identified. We refer to the set of CIs that will need to be modified for the change to be complete ias the
\textit{change set}.

@@

\begin{figure*}[!t]
\centering
\includegraphics[width=15cm]{graphics/constructingmodel.png}
\caption{Building the Bayesian Network}
\label{fig:process}
\end{figure*}

Bayesian Networks, first introduced by Pearl~\cite{pearl1988probabilistic} are a type of probabilistic graphical models. Each node in the network
represents a random
variable, and nodes are connected through edges which represent probabilistic dependencies~\cite{benbayesian}. A Bayesian network is a Directed
Acyclic Graph (DAG), and is essentially defined by two aspects: the structure of the graph and the Conditional Probability Tables (CPTs) probabilistically
relating connected nodes. In our work, a node in the Bayesian network would be a CI which can take values {``true'',``false''} to donate the probability of it
changing or not. The CPTs would essentially give the probability of one CI changing given its parents' values.

Given a set of data consisting of observations for our variable space, machine learning techniques can be used to estimate the network. If the structure of the
network is known, and we have full observability (i.e we can observe values for each variable), then learning the CPTs can be achieved through
Maximum-likelihood estimation (MLE)~\cite{benbayesian}. MLE tries to maximize the log likelihood of the the training data which can simply be done by
calculating the frequency of occurrence of the values of each variable given its parents. If the structure is known, but we only have partial observability
(i.e., missing data), then we can use a variation of MLE which is the Expectation Maximization algorithm~\cite{dempster1977maximum}. If the structure is not
known in
the first place, then searching through the model space, and calculating network scores for each produced network is necessary. The simplest algorithm for this
is a Greedy search algorithm~\cite{friedman1996learning}.

While searching for the best network that fits the training data, scores are needed for the produced networks in order to know which one is the best one. There
are several types of scores that have been proposed in the literature. Two of the commonly used scores are the Minimum Description Length (MDL)
score~\cite{rissanen1986stochastic} and the BDe score~\cite{heckerman1995learning}. The MDL score is based on finding the length of data required to store a
network. It tries to balance the complexity of the induced network with the degree of accuracy of which the network represents the frequencies in the training
data. Therefore, the lower the MDL score of a network, the better it is. The BDe score is proportional to the posterior probability of each network structure
being scored, given the data. Accordingly, we are searching for the network that maximizes this probability. For large data sets, both metrics score
structures similarly~\cite{friedman1996learning}.

@@

There are many machine learning and Bayesian analysis tools available. In this work, we use three of these tools which we briefly describe
below.

\subsubsection{WEKA}

WEKA~\cite{hall2009weka} is a Java toolbox for machine learning. It provides different algorithms for classification and clustering as well as other machine
learning problems. It has Bayesian learning techniques for classification problems. Additionally, it provides explorations tools for Bayesian
Networks, and allows the user to learn the structure and CPTs of a Bayesian Network. IN terms of learning CPTs, one of the algorithms WEKA implements is
the SimpleEstimator algorithm~\cite{witten2005data} implemented which we use in our work. This simply estimates the probability values on edges based on
the frequency values of each variable given its parents in the training set.

\subsubsection{Banjo}
Banjo~\cite{banjotool} is a tool written in Java which infers the structure of a Bayesian network given training data. It has two different search algorithms:
Greedy and Simulated Annealing. Banjo uses the BDe metric to compute a network's score, and produces the highest scoring network found. We used the Greedy
search algorithm in our experiments.

@@

JavaBayes~\cite{javabayestool} is a Java tool for calculating marginal probabilities and expectations in a Bayesian network. In a network, nodes can be set to
be observed, and then
the posterior probability of the remaining nodes can be calculated using Bayes theorem. This is the functionality we need here since the CI that will be
changed will be considered as observed to be ``true'', and then we need to calculate the posterior of the remaining CIs given this observation.

\section{Constructing the Bayesian Network}
\label{sec:constructing}

Before constructing the Bayesian Network, we first have to process and prepare the data we have. In order to construct a Bayesian network to use for
predictions there are two steps involved. First, determining the structure of the actual network, and then estimating the Conditional Probability Tables
(CPT). Section~\ref{sec:data} first explains the data set available to us, and the data preprocessing involved before building the network.
Section~\ref{sec:structure} then explains the different techniques we experimented with to build the network structure. Section~\ref{sec:cpt} shows the last
step to build the network which is estimating the CPTs. Figure~\ref{fig:process} shows this overall process.

@@

\subsubsection*{Data Size}
The data set available to us in the sample CMDB we have is from three years, and has 7,999 distinct CIs, and 27,305 change orders. This amount of data
was infeasible to work with as no tool could handle such a large amount of variables in a Bayesian network. For the purposes of this project, which is mainly to
experiment with Bayesian techniques for change set prediction, it is sufficient to choose a small representative subset of the data. Accordingly, in order to be
able to test things properly, we used observations from three months data from January 1, 2008 to March 31, 2008 to build the model. However, even in such a
short period, there were already 2,841 distinct CIs appearing and 2,229 observations (i.e. change orders). Therefore, we needed to perform further data
preprocessing.

First, we removed all CIs that have changed less than 12 times within this time frame  (i.e were associated with less than 12 different change orders in our
training observations). There is no particular reason for choosing 12 as a cut off. It simply gave a feasible data set to deal with. This yielded 120 distinct
CIs. Then, in order to slightly increase our variable space to include other related CIs, we found all the parent CIs related to these 120 CIs from the CMDB
perspective. Adding the related parent CIs, we now had a set of 241 CIs. However, some of the added parent CIs may not be in the original set of 2,841 CIs.
Accordingly, we just kept CIs that appeared more than 12 times or were in the set of related CIs. This provided us our final data set of 170 CIs which we use
throughout our models for fair comparison. Additionally, filtering out CIs meant filtering out some of the change orders that did not have any CIs satisfying
our criteria (i.e. all the CIs that changed in them were outside our 170 CI set). This led to us having 1,305 change orders instead of 2,229 which was more
manageable set. Accordingly, our training data set consisted of 170 variables (CIs), and 1,305 observations (change sets).

@@

In the CMDB, a Change Order has several fields including the requester, the assignee, the start date of the change, the description and the change set field
(called the `Configuration items' fields). Of the fields in a change order, we use only the change set field. The advantage of only using this one field is that
change sets are easy to extract and easy to understand. In terms of an observation, CIs were either marked as ``true'' or ``false'' indicating whether they
have changed or not. Each observation initially consisted of all the CIs in our data set marked as ``false''. For each change order in the training set, we
would
mark each CI appearing in the ``Configuration Items'' field as true, while all those not appearing were left as ``false''. Therefore, the set of training
data consisted of an observation entry for each change order with the CIs marked as ``true'' or ``false'' accordingly.

\subsection{Network Structure}
\label{sec:structure}
There were essentially three different ways in which we could estimate the structure of the network. We, thus, built a Bayesian Network using each technique to
be able to compare their performance.

\subsubsection{Method 1 (Banjo)}
\label{sec:model1}

For the first method, we used Banjo to learn the structure of the Bayesian network from the training data. We use the Greedy searching algorithm implemented
there. Banjo produced five different top-scoring networks (having the same score), and we simply chose one of them as the network to be used.

\subsubsection{Method 2 (WEKA Classification)}

Although our research problem as currently defined is not really a classification problem, we want to examine how the network built by the BayesNet classifier
in WEKA will perform. Accordingly, we estimated the structure of the network using the K2 structure learning algorithm implemented in WEKA which is also a
greedy
search based learning algorithm. The default maximum number of parents allowed for any node is one (a Naive Bayesian Network). However, we used a maximum of
five to allow the nodes to be related since there is really no class label in our data set. We could not experiment with other techniques in WEKA because they
do not allow multiple parents.

\subsubsection{Method 3 (CMDB Relationships)}

For the third method, we used the relations existing in the CMDB to infer the structure of the network. We used two
variations for this. The first was to follow the direction of the relationship edges in the CMDB. That is, if there is a relationship edge in the CMDB from A to
B, we will place an
edge in the Bayesian network from A to B. The second was to reverse the direction of the relationship edges in the CMDB. That is,
if there is a relationship edge from A to B in the CMDB, we will place an edge in the Bayesian network from B to A. There was one problem, however, with
the generated networks. Two of the CIs had more than 15 parents to them which means that their CPTs will be intractable to compute. For those two CIs, we simply
removed all their parents to make the computation tractable.

@@

After the structure was determined, we used the SimpleEstimator algorithm~\cite{witten2005data} built in WEKA to calculate the CPTs for all networks. After
setting the data set to be the observations in our training set (an ARFF file), WEKA learned the CPTs for each of the
networks above. This complete network was then saved as an BIF XML file to be used by JavaBayes for general inference as will be explained in the next section.

Additionally, we experimented with the missing data feature in WEKA. WEKA allows the user to specify missing values
for any variable by simply putting a '?' instead of its value. Therefore, instead of putting the CIs that did not appear in a change order as false, we put a
'?' in their place. This seemed closer to practice, because in reality, we are not sure whether this CI actually changed or not. When using Banjo or the CMDB
relationships to estimate the structure of the network, the missing values were only inserted when learning the CPTs. When using WEKA, the missing values were
inserted both to learn the network, and to estimate the CPTs.

\section{Experiment Setup}
\label{sec:setup}

This section explains how we tested the predictions of each of the networks in a way that simulates how an analyst would use the tool in reality. We first
explain how we simulate this behavior, and then explain how we evaluate the predicted CIs. Algorithm~\ref{alg:simulation} summarizes these steps taken to
predict a change order.

\begin{algorithm}[tb]
   \caption{Generating Predictions using the Bayesian Network}
   \label{alg:simulation}
\begin{algorithmic}
   \STATE {\bfseries Input:} $changeOrder$
    \STATE {\bfseries Input:} $BayesianNetwork$
\STATE{\textbf{Input:}} $networkCIs$ 
    \STATE {\bfseries Input:} $threshold$
    \STATE Set $initialCI$ = first CI in changeOrder
    \STATE Set $occurredset$ = $changeOrder$ - $initialCI$
    \STATE Initialize $observedSet$ = \{$initialCI$\}
    \STATE Initialize $predictedSet$ = \{\}
   \REPEAT  
    \STATE Initialize $newPredictions$ = \{\}
 \IF{$initialCI$ not in $networkCIs$} 
\STATE return 
\ENDIF
    \FOR{$node$ in $BayesianNetwork$}
    \STATE $posterior =$ query $node$ in $BayesianNetwork$ 
    \IF{$posterior > threshold$ \textbf{and} $node$ \textbf{not in} $predicted set$} 
      \STATE Add $node$ to $newPredictions$   
    \ENDIF
    \ENDFOR

  \FOR{$prediction$ in $newPredictions$}
    \IF{$prediction$ in  $occurredSet$}
      \STATE Add $prediction$ to $observedSet$
    \ENDIF
    \ENDFOR
\STATE Add $newPredictions$ to $predictedSet$

   \UNTIL{$newPredictions$ is \{\}}
\newline
\STATE $recall=  \frac{predictedSet\cap occurredSet}{occurredSet}$
\newline
\STATE $precision = \frac{predictedSet\cap occurredSet}{predictedSet} $
\end{algorithmic}
\end{algorithm}

@@

The change set detection process provided is an iterative, collaborative process between the tool and the analyst. When the tool suggests CIs, the analyst can
accept or reject these CIs (i.e add them to the change set or not). Based on the CIs added to the change set, the analyst can ask the tool for more
suggestions. This is along the lines of ``Now, that I am also going to change these CIs, what else do I need to change?''. This process is explained in more
details by Nadi et al.~\cite{nadi2010}.

To simulate this process for our experiments, we do the following. At this point, we have the Bayesian network ready, and we would like to perform Inference.
More formally, given that a CI will change (our observation), we want to update our beliefs that the other CIs in the network might change as well.
Unfortunately, neither of the two tools previously used provide a general inference engine. We, therefore, use JavaBayes in this step since it accepts
the same XML BIFF format produced by WEKA. We use a one month testing set where we try to predict all the change orders in April 2008. There was a total of
883 change orders in that month which we use in our testing set.

For each change order, we would take the first CI recorded as the initial CI to change, then we would set that as an observed node, and update
the beliefs of all the nodes (CIs) in the network using JavaBayes. We are basically checking the posterior probability of each CI being ``true'' given our
observation. We used different cutoff thresholds for the posterior probabilities we would consider. We would then loop on all the updated CIs, and add those
that match our threshold criteria to the predicted change set. For example a threshold of 0.2 means that only nodes that have a posterior probability greater
than 0.2 of having the value ``true'' will be returned. To simulate a real life scenario, we then checked which of these predicted CIs actually lies in the
target change set we are trying to predict. This is similar to an analyst accepting CIs into their change set. These common CIs would then also be marked as
observed so that we can predict what else will need to change given that these CIs are also changing. Again, we would calculate the posterior probability given
the new observations, and continue doing so until there are no more common CIs. All the CIs that match the threshold criteria (whether accepted by the analyst
or not) are part of the predicted set. These steps (shown in Algorithm~\ref{alg:simulation} would be repeated for each change order. Note that if the initial
CI is not in the list of CIs we used for training, then the predicted set will be empty since we have not seen this CI before.

@@

We need a way to evaluate the predicted CIs. The recall and precision measures from Information Retrieval are appropriate for this type of evaluation.
Recall measures the proportion of correct CIs retrieved by the system, while precision measures the proportion of suggested CIs that are correct~\cite{van79}.

Similar to Hassan et. al~\cite{hassan2004predicting}, we define the \textit{Predicted Set} (P) as the set of all predicted. We define the \textit{Occurred
Set} (O) as the CIs remaining in the change set after excluding the Initial CI
provided by the analyst (i.e Change Set - Initial CI). The intersection of the predicted set and the occurred set, called $PO$, is the common CIs in both sets.
For each constructed change set, we then calculate the recall and precision values for the predictions according to the following
definitions~\cite{hassan2004predicting}:

@@

\begin{table*}[h!t!]
\centering
\resizebox{\textwidth }{!}{
\begin{tabular}{ c | c |c ||c | c || c |c ||c | c || c |c ||c| c||c|c||c|c}
\textbf{Threshold} 
& \multicolumn{2}{|c||}{\textbf{Banjo}} 
& \multicolumn{2}{|c||}{\textbf{Banjo Missing}}
&\multicolumn{2}{|c||}{\textbf{WEKA}}
&\multicolumn{2}{|c||}{\textbf{WEKA Missing}}
&\multicolumn{2}{|c||}{\textbf{CMDB}}
&\multicolumn{2}{|c||}{\textbf{CMDB Missing}}
&\multicolumn{2}{|c||}{\textbf{Reversed CMDB}} 
&\multicolumn{2}{|c}{\textbf{Reversed CMDB Missing}}\\
\hline
 & \textbf{Recall} 
& \textbf{Precision}
& \textbf{Recall} 
& \textbf{Precision}
& \textbf{Recall}
& \textbf{Precision}
& \textbf{Recall} 
& \textbf{Precision}
& \textbf{Recall} 
& \textbf{Precision}
& \textbf{Recall} 
& \textbf{Precision}
& \textbf{Recall} 
& \textbf{Precision}
& \textbf{Recall} 
& \textbf{Precision}\\
\hline\hline
0&0.715&0.473&0.715&0.473&0.715&0.473&0.715&0.473&0.715&0.473&0.715&0.473&0.715&0.473&0.715&0.473\\
\hline
0.1&0.649&0.493&0.715&0.473&0.707&0.501&0.715&0.473&0.569&0.470&0.715&0.473&0.583&0.471&0.715&0.473\\
\hline
0.2&0.643&0.501&0.715&0.473&0.704&0.509&0.715&0.473&0.568&0.470&0.715&0.473&0.580&0.470&0.715&0.473\\
\hline
0.3&0.634&0.510&0.715&0.473&0.700&0.519&0.715&0.473&0.568&0.469&0.715&0.473&0.569&0.469&0.715&0.473\\
\hline
0.4&0.624&0.511&0.715&0.473&0.696&0.524&0.715&0.473&0.568&0.469&0.715&0.473&0.568&0.469&0.715&0.473\\
\hline
0.5&0.610&0.500&0.715&0.473&0.682&0.528&0.715&0.473&0.568&0.469&0.715&0.473&0.568&0.469&0.715&0.473\\
\hline
0.6&0.607&0.500&0.715&0.473&0.674&0.529&0.715&0.473&0.567&0.469&0.715&0.473&0.568&0.469&0.715&0.473\\
\hline
0.7&0.595&0.499&0.715&0.473&0.646&0.526&0.715&0.473&0.567&0.469&0.715&0.473&0.568&0.469&0.715&0.473\\
\hline
0.8&0.595&0.499&0.715&0.473&0.627&0.526&0.715&0.473&0.567&0.469&0.715&0.473&0.567&0.469&0.714&0.473\\
\hline
0.9&0.574&0.475&0.715&0.473&0.588&0.489&0.567&0.469&0.567&0.469&0.715&0.473&0.567&0.469&0.712&0.473\\
\hline
1&0.567&1.000&0.567&1.000&0.567&1.000&0.567&1.000&0.567&1.000&0.567&1.000&0.567&1.000&0.567&1.000\\
\end{tabular}
}
\centering
\caption{Summary of Results}
\label{tab:results}
\end{table*}
